The Role of Video Data Sharing in Advancing Education Policy and Practice

### 2015-05-08 2015-06-01-nsf-ed-workshop/paper.tex
draft of intro

\abstract{
Video captures the complexity, richness, and diversity of behavior unlike any other measure. 
As a result, large numbers of people who study teaching and learning or seek to improve it employ video. 
Video documents itself to a large degree, and therefore has significant potential for data re-use.
This potential remains largely unrealized because videos are rarely shared.
Video often contains information about personal identities, so considerations about research ethics pose
challenges to sharing. 
The relatively large size of video files and diversity of formats and software tools for coding pose technical challenges. 
In this thought paper, we will describe how the web-based Databrary data library has overcome the most significant barriers to sharing video within the developmental sciences community, including solutions to maintaining participant privacy, data tagging, and data management.
The Databrary model and system provides insights about how video and other identifiable data collected in the context of education research might be shared.
Greater data sharing will advance education policy and improve practice.
}

@@

\section{Introduction}
Scientific discovery and the application of discovery to human needs such as improved educational policy and practice depends on the free flow of information. 
According to the NIH Data Sharing Policy (NIH 1), open data sharing is essential for translating research findings into improvements in human health. 
Open data sharing promotes greater transparency and peer oversight, supports inquiry into data collection methods and measurement, stimulates new work, encourages diversity of analysis and opinion, allows for testing of new or alternative hypotheses, facilitates education of new researchers, permits integrative analyses by creating new datasets from multiple sources, allows future research to build on earlier efforts, and enables exploration of topics not envisioned by the original investigators. 
In short, open data sharing leads to new uses and scientific insights and thus increases the impact of public investments in research.
Open data sharing is a scientific imperative and common practice in many areas of biomedical, physical, and earth sciences where it has accelerated the pace of discovery (NIH 2–5). Unfortunately, research on behavioral development remains shrouded in a culture of isolation. 
Rather than providing direct access to raw data, researchers share interpretations of distilled data through publications and presentations5.

Reliance on video impedes data sharing in developmental science. 
Most developmental researchers rely on video because it is cheap, easy to use, and faithfully renders the richness and complexity of behavior6. 
However, researchers rarely share raw video data—even within labs and among collaborators. 
After publication, the rich, largely under-used videos molder quietly on a shelf. This situation is ironic because video data are conducive to sharing; the behaviors recorded are directly available for peer inspection and the complexity of video data makes it amenable to mining for new information. 
Certainly, sharing digital video poses technical problems and challenges to protection of participant privacy. But, failure to overcome the obstacles to sharing has impeded transparency and peer oversight, limited the generalizability of findings, slowed progress, and diminished the impact of Federal funding. The Databrary project aims to increase scientific transparency and speed insights into behavioral development by building the community practices and infrastructure for open video data sharing. 
The contribution of a particular dataset will no longer depend on the private activities of one researcher, but will instead benefit from the imagination of many researchers with different viewpoints.

Video is a uniquely rich medium for capturing in real time the interactions that occur in classroom, laboratory, and informal learning settings. 
Most researchers in the developmental and learning sciences collect video as raw data, but ethical and technical barriers to sharing video have made open data sharing uncommon. 
While personally identifying information can be removed from text-based data, videos may contain faces, voices, names spoken aloud, and sometimes views of the homes of research participants. 
These cannot be removed without reducing the information content. Further, the collection of video or other identifiable or sensitive information requires approval by a research ethics board and informed consent from the participants. 
The consent process formalizes a promise by the research team that a participant's identity will not be disclosed. Thus, researchers risk violating participants’ privacy if digital images are viewed or released to others without authorization.

In this paper, we describe the promise of video data sharing, the challenges it poses, and the solutions that the Databrary team has created to address them.
In developing technology and policies that enable the safe and secure sharing of video and other identifiable research data, Databrary illustrates how researchers interested in advancing educational policy and practice through widespread video data sharing might achieve their ambitious aims.

@@

@@

\section{The Challenges of Video Data Sharing}

@@

\subsubsection{Online Video Sharing}

\begin{itemize}
\item Size of files
\item Storage/streaming
\item Formats/transcoding
\item Preservation
\end{itemize}

\subsubsection{Tools for Coding Video}
After videos are recorded, human coders evaluate videos and apply text-based labels or numeric codes to videos.
Coders use a wide-range of commercial and academic software tools: Transana, StudioCode, V-code, MaxQDA, Noldus Observer, Mangold Interact, and Datavyu, among others.
Some use spreadsheets or pencil and paper.
The tools have no standard format for data storage or export.
Some offer web/cloud based storage (e.g., Transana, V-code), but otherwise coded data may not be easily shared within tools, much less between them.
With few exceptions, machine learning has not been applied to video data coding.
\subsection{Ethics and Privacy}
Sharing video recordings poses a unique challenge to existing data sharing policies because videos contain personally identifying information—specifically participants’ faces and voices and often the
insides of their homes and classrooms. 
Sharing personally identifiable information puts research participants and others depicted in recordings at increased risk for loss of privacy. 
At the same time, blurring or altering original recordings to hide identities undermines or eliminates
their value to other researchers. 
Often, participants’ faces and voices produce the behaviors of interest.
A challenge for future education research is to find ways to preserve the value of video for reuse by others while protecting the privacy of research participants.
Databrary has developed an access model described below that provides an example of how this might be accomplished.
\subsection{Data Management}
A primary impediment to sharing is researchers’ lack of time to find, label, clean, organize, and copy their files into formats that can be used and understood by others15. 
Even scientists committed to data sharing find it difficult to do. Data management systems that make sharing convenient, reliable—and optimally automatic29—can help.

>>>>

\abstract{
Video captures the complexity, richness, and diversity of behavior unlike any other measure. 
As a result, large numbers of people who study teaching and learning or seek to improve it employ video. 
Video documents itself to a large degree, and therefore has significant potential for reuse by others.
This potential remains largely unrealized because videos are rarely shared.
Video often contains information about personal identities, so considerations about research ethics pose
challenges to sharing. 
The relatively large size of video files and diversity of formats and software tools for coding pose technical challenges. 
In this thought paper, we will describe how the Databrary data library has overcome the most significant barriers to sharing video within the developmental sciences community, including solutions to maintaining participant privacy, data tagging, and data management.
Databrary suggests ways that video and other identifiable data collected in the context of education research might be shared.
Widespread open sharing of high value, high impact data will advance education policy and improve practice.
}

@@

\section{Introduction}

The application of insights from research to essential human needs such as improved educational policy and practice depends on the free flow of information. 
Open data sharing promotes greater transparency and peer oversight, supports inquiry into data collection methods and measurement, stimulates new work, encourages diversity of analysis and opinion, allows for testing of new or alternative hypotheses, facilitates the education of new researchers, permits integrative analyses by creating new datasets from multiple sources, allows future research to build on earlier efforts, and enables exploration of topics not envisioned by the original investigators. 
Open data sharing increases the impact of public investments in research and leads to better and more effective public policy.
Open data sharing is a scientific imperative and common practice in many areas of biomedical, physical, and earth sciences where it has accelerated the pace of discovery. 
Unfortunately, research on human development and learning remains shrouded in a culture of isolation. 
Rather than providing direct access to raw data, researchers in the developmental and learning sciences share interpretations of distilled data almost exclusively through publications and presentations.
The path from raw data to finding to conclusion cannot be traced or validated by others, nor can new questions be posed by others building on the same, often expensive-to-collect, raw data.
To realize significant advances in our understanding of teaching and learning and for those insights to have meaningful impact on educational practice and policy, open data sharing must become an established norm, not the rare exception.
Moreover, we must find ways to share the most valuable raw data, the data with the greatest potential for reuse by others, and the data with the greatest potential for yielding meaningful insights into the learning process.

Video is a uniquely rich medium for capturing in real time the interactions that occur in classrooms, muesuems, laboratory, home, and other informal learning settings.
Many researchers and practitioners in the developmental, learning, and education sciences collect video as raw data.
Video recordings capture what teachers say and do and what students say and do in ways that offer the potential to discover when, why, and how learning actually works.
Still, the open sharing of video data is rare, even within labs or among collaborators.
Video data should be conducive to sharing.
Video captures behavior that others can readily inspect.
It captures multiple dimensions of complex real-world learning settings in ways that others can easily build upon. But, technical, ethical, and cultural barriers have made open sharing of video data rare.
Videos come in varied formats and generage large files; these pose conversion and storage challenges.
Personally identifying information can be removed from text-based data.
Videos may contain faces, voices, names spoken aloud, views or voices of non-participants, views of classroom interiors and sometimes views of the homes of research participants. 
These elements cannot be removed without reducing the information content and reuse value to others. 
The collection of video or other identifiable or sensitive information requires approval by a research ethics board.
It also requires informed consent or assent from the participants, and possibly parental, teacher, and school district approval.
Researchers risk violating participants’ privacy if digital images are viewed or released to others without authorization.
For these and other reasons, no culture of widespread open sharing of video data has emerged in the education, learning, and developmental sciences.

The Databrary project has built a digital data library specialized for the open sharing of video data from research in these areas of inquiry.
In this paper, we will describe how Databrary has overcome the most significant barriers to sharing video, including solutions to maintaining participant privacy, data tagging, and data management.
In developing technology and policies that enable the safe and secure sharing of video and other identifiable research data, Databrary suggests ways that video and other identifiable data collected in the context of research on teaching and learning might be shared.
Fostering the open sharing and reuse of high value, high impact data like video promises to advance education policy and improve practice.

@@

More than 1 billion users worldwide uploade more than 300 hours of video to YouTube every minute (https://www.youtube.com/yt/press/statistics.html). Google, YouTube's owner, has built an immense data infrastructure to upload, store, convert, tag, and stream video. The current scale of video collection in developmental and education contexts is undoubtedly smaller, but it is not small. The Gates Foundation-funded Measures of Effective Teaching (MET) Project, hosted at ICPSR, contains X video segments from X classroom settings, representing tens of terabytes of data. In just over a year of operation, the Databrary digital library has collected more than 7,000 individual videos, representing 2,400 hours of recording, featuring more than 1,800 infant, child, and adult participants. Video data is big, and the interest in recording and sharing video for research, education, and policy purposes has grown.

@@

We record teachers in training and teachers in-service for evaluation.
We record classrooms to capture what teachers do and how students respond.
We record visitors and docents in museum settings.
\section{The Challenges of Video Data Sharing}

@@

Video poses two types of technical challenges related to the goal of increasing widespread sharing. One concerns the requirements for storing, sharing, streaming, and preserving large numbers of videos. The other concerns tools for extracting meaningful information from the episodes and behavior videos capture.

\subsubsection{Video Data is Big}

Video often accompanies other data streams (physiological recordings, brain imaging, EMG, motion tracking, eye tracking, verbal transcripts). 
Multiple data streams—including two or more camera views—require tools for synchronization and integration and benefit from multivariate time series analyses. 
Many labs lack the resources to adopt tools for detailed video coding and complex analyses. 
Furthermore, widespread use of video creates a data explosion: The typical developmental lab collects 8-12 hours of video/week 18 in widely varied formats. 
Thus, sharing digital video requires substantial storage capacity, powerful search and streaming tools, and significant computational resources for transcoding videos into common, preservable formats.

\subsubsection{Tools for Coding Video}
After videos are recorded, human coders evaluate videos and apply text-based annotations or numeric codes to videos.
Coders use a wide-range of commercial and academic software tools: Transana, StudioCode, V-code, MaxQDA, Noldus Observer, Mangold Interact, and Datavyu, among others.
Some use spreadsheets or pencil and paper.
The tools have no standard format for data storage or export.
Some offer web/cloud based storage (e.g., Transana, V-code), but otherwise coded data may not be easily shared within tools, much less between them.
With few exceptions, machine learning has not been applied to video data coding.
\subsection{Ethics and Privacy}
Sharing video recordings poses a unique challenge to existing data sharing policies because videos contain personally identifying information—specifically participants’ faces and voices and often the
insides of their homes and classrooms. 
Sharing personally identifiable information puts research participants and others depicted in recordings at increased risk for loss of privacy. 
At the same time, blurring or altering original recordings to hide identities undermines or eliminates
their value to other researchers. 
Often, participants’ faces and voices produce the behaviors of interest.
A challenge for future education research is to find ways to preserve the value of video for reuse by others while protecting the privacy of research participants.
Databrary has developed an access model described below that provides an example of how this might be accomplished.
\subsection{Data Management}
A primary impediment to sharing is researchers’ lack of time to find, label, clean, organize, and copy their files into formats that can be used and understood by others15. 
Even scientists committed to data sharing find it difficult to do. Data management systems that make sharing convenient, reliable—and optimally automatic29—can help.
